
## Examples 

```{r}
data(oven)
ovenc <- oven
ovenc[, c(4:8,10:11)][] <- lapply(ovenc[, c(4:8,10:11)], scale)
moven <- svabu(count ~ pforest | observ + pforest + julian + timeday, ovenc)
summary(moven)
drop1(moven, model="det")
moven2 <- update(moven, . ~ . | . - timeday)
summary(moven2)


data(databu)
## fit BZIP and BP models
m00 <- svabu(Y ~ x1 + x5 | x2 + x5, databu[1:200,])
## print method
m00
## summary: CMLE
summary(m00)
## coef
coef(m00)
coef(m00, model="sta") ## state (abundance)
coef(m00, model="det") ## detection
coef(m00, model="zif") ## zero inflation (this is part of the
'
true state
'
!)
## Not run:
## Diagnostics and model comparison
m01 <- svabu(Y ~ x1 + x5 | x2 + x5, databu[1:200,], zeroinfl=FALSE)
## compare estimates (note, zero inflation is on the logit scale!)
cbind(truth=c(2,-0.8,0.5, 1,2,-0.5, plogis(0.3)),
"B-ZIP"=coef(m00), "B-P"=c(coef(m01), NA))
## fitted
plot(fitted(m00), fitted(m01))
abline(0,1)
## compare models
AIC(m00, m01)
BIC(m00, m01)
logLik(m00)
logLik(m01)
## diagnostic plot
plot(m00)
plot(m01)

# Nonparametric bootstrapping
# - initial values are the estimates
m02 <- bootstrap(m00, B=25)
attr(m02, "bootstrap")
extractBOOT(m02)
summary(m02)
summary(m02, type="cmle")
summary(m02, type="boot")
## vcov
vcov(m02, type="cmle")
vcov(m02, type="boot")
vcov(m02, model="sta")
vcov(m02, model="det")
## confint
confint(m02, type="cmle") ## Wald-type
confint(m02, type="boot") ## quantile based
## parametric bootstrap
simulate(m00, 5)
m03 <- bootstrap(m00, B=5, type="param")
extractBOOT(m03)
summary(m03)
## Model selection
m04 <- svabu(Y ~ x1 + x5 | x2 + x5 + x3, databu[1:200,], phi.boot=0)
m05 <- drop1(m04, model="det")
m05
m06 <- svabu.step(m04, model="det")
summary(m06)
m07 <- update(m04, . ~ . | . - x3)
m07
## Controls
m00$control
getOption("detect.optim.control")
getOption("detect.optim.method")
options("detect.optim.method"="BFGS")
m08 <- svabu(Y ~ x1 + x5 | x2 + x5, databu[1:100,])
m08$control ## but original optim method is retained during model selection and bootstrap
## fitted models can be used to provide initial values
options("detect.optim.method"="Nelder-Mead")
m09 <- svabu(Y ~ x1 + x5 | x2 + x5, databu[1:100,], inits=coef(m08))
## Ovenbirds dataset
data(oven)
ovenc <- oven
ovenc[, c(4:8,10:11)][] <- lapply(ovenc[, c(4:8,10:11)], scale)
moven <- svabu(count ~ pforest | observ + pforest + julian + timeday, ovenc)
summary(moven)
drop1(moven, model="det")
moven2 <- update(moven, . ~ . | . - timeday)
summary(moven2)


```




From: http://dcr.r-forge.r-project.org/qpad/QPAD_SupportingInfo.pdf
4.3.2 Bayesian and frequentist approach to hierarchical modeling
In the next example we consider a Poisson-Log-Normal generalized linear mixed model with a random intercept for routes, which also incorporates detectability related uncertainty. This model uses a global maximization technique called data cloning (Lele et al., 2007, 2010), which takes advantage of Bayesian MCMC techniques for maximum likelihood estimation. The software implementation is described in Solymos (2010). 

### *Note that using a single clone is identical to the Bayesian hierarchical modeling.*

Example GLMM with data cloning

```{r}
library(dcmle) # load dcmle package (loads detect as well)
load_BAM_QPAD(version = 1)

load.module("glm") # load glm module for JAGS

model <- function() {
  for(i in 1:n) {
    Y[i] ~ dpois(lam[i])
    log(lam[i]) <- inprod(X[i, ], beta) + E[gr[i]] + log(A[i] * p[i])
    p[i] <- 1 - exp(-3 * phi[i])
    A[i] <- 3.141593 * tau[i] ^ 2
    log(phi[i]) <- inprod(Z1[i, ], theta01)
    log(tau[i]) <- inprod(Z2[i, ], theta02)
    }
  for(j in 1:m) {
    E[j] ~ dnorm(0, 1 / exp(log.sigma) ^ 2)
    }
  for(k in 1:np) {
    beta[k] ~ dnorm(pr[k], 1)
    }
  log.sigma ~ dnorm(-2, 0.01)
  theta01 ~ dmnorm(theta1, Sigma1)
  theta02 ~ dmnorm(theta2, Sigma2)
  }

bc0 <- with(oven, globalBAMcorrections("OVEN", t = dur, r = dist))
summary(bc0)

bm <- bestmodelBAMspecies("OVEN", type = "BIC")
bc <- with(oven, localBAMcorrections("OVEN", t = dur, r = dist,
    jday = JDAY, tssr = TSSR, tree = pforest, lcc = LCC, model.sra = bm$sra,
    model.edr = bm$edr))
summary(bc)


dat <- list(Y = oven$count, 
            X = model.matrix( ~ pforest + long, oven), 
            np = 3, 
            n = nrow(oven), # every record
            m = length(unique(oven$route)), # site
            gr = dciid(as.integer(as.factor(oven$route))), 
            pr = coef(mod), 
            theta1 = coefBAMspecies("OVEN", bm$sra, bm$edr)$sra, 
            Z1 = model.matrix( ~ JDAY, oven), 
            Sigma1 = solve(vcovBAMspecies("OVEN", bm$sra, bm$edr)$sra), 
            theta2 = coefBAMspecies("OVEN", bm$sra, bm$edr)$edr, 
            Z2 = model.matrix( ~ LCC, oven), 
            Sigma2 = solve(vcovBAMspecies("OVEN", bm$sra, bm$edr)$edr))

dcf <- makeDcFit(model = model, data = dat, params = c("beta", "log.sigma"), multiply = c("n", "m"), unchanged = c("np", "pr", "theta1", "theta2", "Sigma1", "Sigma2"))

cl <- makePSOCKcluster(3) # parallel computing for speed up
K <- c(1, 2) # sequence for the number of clones to use
parLoadModule(cl, "glm") # load glm module for JAGS on workers

dcm <- dcmle(dcf, n.clones = K, n.update = 2000, n.iter = 2000, cl = cl, partype = "parchains")

stopCluster(cl) # close cluster
summary(dcm)
```



