Shenandoah Single vs Multi-pass Sampling
===========================================

Cleaning before entry into R:
* Did text to columns by comma to separate coded notes into columns
* Replaced "MORT" and other non-numeric mortality notes with 7
* Made a 'dead' column that was comprised of notes 7 (mortality) and 10 (whole fish collected)
* Replaced `'` and `"` with blank to avoid problems reading in text columns
* Replaced `ND` (No Data) with `NA` for missing data
* Changed `Species` to `Code` in the fish data headers to be able to merge with the species data (could do in R after import)
* Added primary key
* Removed `#` symbol from GameFish

```{r setup, results = 'hide', message = FALSE, warning = FALSE}
rm(list=ls())

library(data.table) # fast, efficient reading in of big data
library(dplyr)
library(ggplot2)
library(lubridate)

setwd('/Users/Dan/Documents/Research/Stream_Climate_Change/Brook_Trout/singlePassDetection/')

gameFish <- fread("inst/extdata/W_FI_Gamefish.csv", sep = ",", na.strings = c("NA", "ND", "", " "))

gameFish[gameFish == 99] <- NA

siteVisits <- fread("inst/extdata/R_SiteVisits.csv", sep = ",")

siteVisits[siteVisits == 999] <- NA

species <- fread("inst/extdata/W_FI_zdd_FISH_Species.csv", sep = ",", header = TRUE)
upstream <- fread("inst/extdata/shenandoahUpstreamStats.csv", sep = ",")

siteVisits$SiteID <- siteVisits$SITEID
```

```{r merge data frames & reshape to 1 line per site-visit-pass-species-age}

# I love magrittr and pipes!
fish <- gameFish %>%
  group_by(SiteVisit_ID, RUN, Code, AGE) %>%
  summarise(meanTL = mean(TL), count = n()) %>%
  left_join(species, by = 'Code') %>%
  left_join(siteVisits, by = 'SiteVisit_ID')

```

### Fill out RUN, species, and age combinations
This is not efficient at all but I was having trouble getting the columns to fill out properly, so I just brute forced it.

```{r Fill out RUN, species, and age combinations}
fish$siteVisitPass <- paste0(fish$SiteVisit_ID, "--", fish$RUN)

fullGrid <- expand.grid(siteVisitPass = unique(fish$siteVisitPass), Code = unique(fish$Code), AGE = 0:1)

fullGrid$siteVisitPass <- as.character(fullGrid$siteVisitPass)
fullGrid$Code <- as.character(fullGrid$Code)

fullDF <- fullGrid %>%
  left_join(fish, by = c("siteVisitPass", "Code", "AGE"))

fullDF$SiteVisit_ID <- matrix(unlist(strsplit(fullDF$siteVisitPass, split = "--", fixed = TRUE)), ncol = 2, byrow = TRUE)[ , 1]

fullDF$RUN <- matrix(unlist(strsplit(fullDF$siteVisitPass, split = "--", fixed = TRUE)), ncol = 2, byrow = TRUE)[ , 2]

foo <- select(siteVisits, SiteVisit_ID, SITEID, sDate)
fullDF2 <- left_join(fullDF, y = foo, by = c("SiteVisit_ID"))
fullDF2$SiteID <- fullDF2$SITEID.y

allDF <- left_join(fullDF2, upstream, by = 'SiteID')
allDF[is.na(allDF$count), ]$count <- 0 # replace count NA with 0 because now all the combinations should represent site visits and passes that were conducted for all species and ages

allDF$date <- ymd_hms(allDF$sDate.y, tz = "US/Eastern")

```


```{r}

# reduce to the variables of interest for modeling and rename the variables with consistent format

subDF <- select(allDF,  
                        SiteID,
                        SiteVisit_ID,
                        date,
                        Code,
                        RUN,
                        AGE,
                        count,
                        FI_Shocker_Voltage,
                        FEATUREID,
                        Latitude,
                        Longitude,
                        TotDASqKM,
                        ReachSlope,
                        ReachElev,
                        tmaxann_brkt,
                        forest,
                        developed)

setnames(subDF, old = names(subDF), new = c("site",
                                                            "SiteVisit_ID",
                                                            "date",
                          "spp",
                          "pass",
                          "age",
                          "count",
                          "voltage",
                          "FEATUREID",
                          "lat",
                          "lon",
                          "drainageArea",
                          "slope",
                          "elev",
                          "tmaxAnn",
                          "forest",
                          "developed"))

```

### remove sites missing covariate data and filter to brook trout

```{r}
bktDF <- subDF %>%
  filter(!is.na(lat)) %>%
  filter(spp == "BKT")

# check 
subDF %>%
  group_by(spp) %>%
  summarise(siteVisitsPasses = n(), sum(count))

gameFish %>%
  group_by(Code) %>%
  summarise(siteVisitsPasses = n())

```

### Prep climate data
Use the same flow and temperature for all sites
```{r}
load("data/localShenDaymet.RData")

climate <- fread('inst/extdata/Shen_climate_data.csv', sep=',')

setnames(climate, names(climate), c("year", "yday", "tmax", "tmin", "precip", "siteID"))

climate$day.year <- paste0(climate$yday, '-', climate$year)
climate$date <- parse_date_time(climate$day.year, "%j%Y", tz = "US/Eastern")

# move this function to functions in R/
getSeason <- function(DATES) {
    WS <- as.Date("2012-12-15", format = "%Y-%m-%d") # Winter Solstice
    SE <- as.Date("2012-3-15",  format = "%Y-%m-%d") # Spring Equinox
    SS <- as.Date("2012-6-15",  format = "%Y-%m-%d") # Summer Solstice
    FE <- as.Date("2012-9-15",  format = "%Y-%m-%d") # Fall Equinox

    # Convert dates from any year to 2012 dates
    d <- as.Date(strftime(DATES, format="2012-%m-%d"))

    ifelse (d >= WS | d < SE, "Winter",
      ifelse (d >= SE & d < SS, "Spring",
        ifelse (d >= SS & d < FE, "Summer", "Fall")))
}

climate$season <- getSeason(climate$date)
annual.climate <- climate %>%
  group_by(year, season) %>%
  summarise(mean.tmax = mean(tmax), mean.precip = mean(precip))

write.table(annual.climate, file = 'inst/extdata/annual-climate.csv', row.names = FALSE, sep = ",")

masterData$season <- getSeason(masterData$date)

annual.site.climate <- masterData %>%
  group_by(site, year, season) %>%
  summarise(mean.tmean = mean(airTemp), mean.precip = mean(prcp))

save(annual.site.climate, file = 'data/annual-site-climate.RData')
```










